Chemically Self-Consistant Modeling of the Globular Cluster NGC 2808 and its Effects on the Inferred Helium abundance of Multiple Stellar Populations.

### 2023-12-10 src/sections/fidanka.tex
Fixed broken figure labels

If \fidanka were to simply apply the same algorithm to the verticalized CMD
then the resulting fiducial line would likely be a re-extraction of the initial
fiducial line guess. To avoid this, we take a more robust, number density based
approach, which considers the distribution of stars in both color and magnitude
space simultaneously. For each star in the CMD we first using an
\texttt{introselect} partitioning algorithm to select the 50 nearest stars in
F814W vs. F275W-F814W space. To account for the case where the star is at an
extreme edge of the CMD, those 50 stars include the star itself (such that we
really select 49 stars + 1). We use
\texttt{qhull}\footnote{https://www.qhull.com}\citep{Barber1996} to calculate
the convex hull of those 50 points. The number density at each star then is
defined as $50/A_{hull}$, where $A_{hull}$ is the area of the convex hull.
Because we use a fixed number of points per star, and a partitioning algorithm
as opposed to a sorting algorithm, this method scales like $\mathcal{O}(n)$,
where n is the number of stars in the CMD. This method also intrinsically
weights the density of of each star equally as the counting statistics per bin
are uniform. We are left with a CMD where each star has a defined number
density (Figure \ref{fig:densityMapDemo}).

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{src/figures/notebookFigures/DensityMapDemo.png}
	\label{fig:densityMapDemo}
	\caption{Density map demo showing density estimate over different parts of
	the evolutionary sequence. The left panel shows the density map over the
	entire evolutionary sequence, while the middle panel shows the density map
	over the main sequence and the right most panel shows the density map over
	the RGB. Figures in the top row are the raw CMD, while figures in the
	bottom row are colored by the density map.}
\end{figure*}

\fidanka can now exploit this density map to fit a better fiducial line to the
data, as the density map is far more robust to outliers. There are multiple
algorithms we implement to fit the fiducial line to the color-density profile
in each magnitude bin (Figure \ref{fig:densityBinsDemo}); they are explained in
more detail in the \fidanka documentation. However, of most relevance here is
the Bayesian Gaussian Mixture Modeling (BGMM) method. BGMM is a clustering
algorithm which, for some fixed number of n-dimensional Gaussian distributions,
$K$, determines the mean, covariance, and mixing probability (somewhat
analogous to amplitude) of each $k^{th}$ distribution, such that the local
lower bound of the likelyhood of each star belonging strongly to a single
distribution is maximized.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{src/figures/notebookFigures/DensityBinsDemo.png}
	\label{fig:densityBinsDemo}
	\caption{CMD where points are colored by density. Lines show the
	density-color profile in each magnitude bin. In this figure adaptive
	binning targeted 1000 stars per bin}
\end{figure}

>>>>

If \fidanka were to simply apply the same algorithm to the verticalized CMD
then the resulting fiducial line would likely be a re-extraction of the initial
fiducial line guess. To avoid this, we take a more robust, number density based
approach, which considers the distribution of stars in both color and magnitude
space simultaneously. For each star in the CMD we first using an
\texttt{introselect} partitioning algorithm to select the 50 nearest stars in
F814W vs. F275W-F814W space. To account for the case where the star is at an
extreme edge of the CMD, those 50 stars include the star itself (such that we
really select 49 stars + 1). We use
\texttt{qhull}\footnote{https://www.qhull.com}\citep{Barber1996} to calculate
the convex hull of those 50 points. The number density at each star then is
defined as $50/A_{hull}$, where $A_{hull}$ is the area of the convex hull.
Because we use a fixed number of points per star, and a partitioning algorithm
as opposed to a sorting algorithm, this method scales like $\mathcal{O}(n)$,
where n is the number of stars in the CMD. This method also intrinsically
weights the density of of each star equally as the counting statistics per bin
are uniform. We are left with a CMD where each star has a defined number
density (Figure \ref{densityMapDemo}).

\begin{figure*}
	\centering
	\includegraphics[width=0.9\textwidth]{src/figures/notebookFigures/DensityMapDemo.png}
	\caption{Density map demo showing density estimate over different parts of
	the evolutionary sequence. The left panel shows the density map over the
	entire evolutionary sequence, while the middle panel shows the density map
	over the main sequence and the right most panel shows the density map over
	the RGB. Figures in the top row are the raw CMD, while figures in the
	bottom row are colored by the density map.}
	\label{densityMapDemo}
\end{figure*}

\fidanka can now exploit this density map to fit a better fiducial line to the
data, as the density map is far more robust to outliers. There are multiple
algorithms we implement to fit the fiducial line to the color-density profile
in each magnitude bin (Figure \ref{densityBinsDemo}); they are explained in
more detail in the \fidanka documentation. However, of most relevance here is
the Bayesian Gaussian Mixture Modeling (BGMM) method. BGMM is a clustering
algorithm which, for some fixed number of n-dimensional Gaussian distributions,
$K$, determines the mean, covariance, and mixing probability (somewhat
analogous to amplitude) of each $k^{th}$ distribution, such that the local
lower bound of the likelyhood of each star belonging strongly to a single
distribution is maximized.

\begin{figure}
	\centering
	\includegraphics[width=0.45\textwidth]{src/figures/notebookFigures/DensityBinsDemo.png}
	\caption{CMD where points are colored by density. Lines show the
	density-color profile in each magnitude bin. In this figure adaptive
	binning targeted 1000 stars per bin}
	\label{densityBinsDemo}
\end{figure}
